---
title: "SOM Tables"
author: "Tyler Clavelle"
date: "August 25, 2015"
output: word_document
---

```{r,include=FALSE}
library(knitr)
library(dplyr)

### Load data -----------------------------------------------------

## NEI
nei<-read.csv('/Results/5.1 global demand spcat phi/Data/UnLumped Projection DataAll Stocks Profit Upsides From NEIs.csv',stringsAsFactors=F)

nei$Percent.of.Total.Profit.Upside.from.NEIs<-format(round(nei$Percent.of.Total.Profit.Upside.from.NEIs,2),nsmall=2)

colnames(nei)<-c('X','Country','Policy','Scenario','Identification Level','Profit Upside Relative to BAU from NEIs','Total Profit Upside','Percent of Total Profit Upside from NEIs')

nei$Scenario<-'Cons. Concern'

## Costs

# Cost to revenue ratios
costratio<-read.csv('/Users/Tyler/Documents/Sustainable Fisheries Group/SFG 2014/Global Fishery Potential/Results/5.1 global demand spcat phi/Data/Cost_Revenue_Ratios_ISSCAAP.csv',stringsAsFactors=F)

# Cost per ton
costperton<-read.csv('/Users/Tyler/Documents/Sustainable Fisheries Group/SFG 2014/Global Fishery Potential/Results/5.1 global demand spcat phi/Data/CostPerTon_Ratios_ISSCAAP.csv',stringsAsFactors=F)

cost<-inner_join(costratio[,c('ISSCAAP.Group','Median','Mean')],costperton[,c('ISSCAAP.Group','Median','Mean')],by=c('ISSCAAP.Group'))

colnames(cost)<-c('ISSCAAP Category','Median Ratio','Mean Ratio','Median $/Ton','Mean $/Ton')

## Global results for fig 3
fig3<-read.csv('/Users/Tyler/Documents/Sustainable Fisheries Group/SFG 2014/Global Fishery Potential/Results/5.1 global demand spcat phi/Data/Fig3data_Profit2050.csv',stringsAsFactors=F)

fig3$Policy<-c('Business As Usual S1','Business As Usual S2','RBFM S1','RBFM S2','FMSY S1','FMSY S2','Today')

fig3<-fig3[,c('Policy','Bio','Profit','Catch')]

fig3$Bio<-format(round(fig3$Bio,2),nsmall=2)
fig3$Profit<-format(round(fig3$Profit,2),nsmall=2)
fig3$Catch<-format(round(fig3$Catch,2),nsmall=2)

colnames(fig3)<-c('Policy','Biomass (MMT)','Annual Profit ($ Billions)','Harvest (MMT)')

## RAM stock list
# ram<-read.csv('/Users/Tyler/Documents/Sustainable Fisheries Group/SFG 2014/Global Fishery Potential/Results/5.1 global demand spcat phi/Data/Ram Stock List.csv',stringsAsFactors=F)
# 
# ram<-ram[,c('IdOrig','CommName','Country','RegionFAO')]
# 
# colnames(ram)<-c('Stock ID','Stock Name','Assigned Country','FAO Region')

## Species specific phi values


```


**Table S5**  
Fishing costs as a fraction of revenue and per-ton catch. Results suggest that profitability varies considerably across ISSCAAP categories
```{r,echo=FALSE,results='asis'}
kable(cost)
```

**Table S6**  
Global results for all policies (BAU, RBFM, and FMSY) and scenarios (S1 and S2) in 2050 as presented in Figure 3 of the main text  
```{r,echo=F,results='asis'}
kable(fig3)
```

**Table S7**  
Percent of upside in profit from NEI fisheries by Country for the RBFM policy applied to stocks of conservation concern
```{r,echo=FALSE,results='asis'}
kable(nei[,c('Country','Scenario','Policy','Percent of Total Profit Upside from NEIs')])
```

**Table S8**  
Fisheries included from the RAM Legacy Stock Assessment Database. Upsides for fisheries listed as "Multinational" were distributed to countries identified as participating in that fishery using FAO landings data (SOM)
```{r,echo=F,results='asis'}
# kable(ram)
```


