---
title: "SOM Tables"
author: "Tyler Clavelle"
date: "August 25, 2015"
output: word_document
---

```{r,include=FALSE}
library(knitr)
library(dplyr)

### Load data -----------------------------------------------------

Batch<-

## NEI
nei<-read.csv('Results/6.0 global demand common phi/Data/UnLumped Projection DataAll Stocks Profit Upsides From NEIs.csv',stringsAsFactors=F)

# nei$Percent.of.Total.Profit.Upside.from.NEIs<-format(round(nei$Percent.of.Total.Profit.Upside.from.NEIs,2),nsmall=2)

nei<-nei %>%
  select(X,Country,Policy,Scenario,Identification.Level,Percent.of.Total.Country.Profits.from.NEIs)

nei$Percent.of.Total.Country.Profits.from.NEIs<-format(round(nei$Percent.of.Total.Country.Profits.from.NEIs,2),nsmall=2)

colnames(nei)<-c('X','Country','Policy','Scenario','Identification Level','Percent of Country Profits from NEIs')

nei$Scenario<-'Cons. Concern'

## Costs

# Cost to revenue ratios
costratio<-read.csv('Results/6.0 global demand common phi/Data/Cost_Revenue_Ratios_ISSCAAP.csv',stringsAsFactors=F)

# Cost per ton
costperton<-read.csv('Results/6.0 global demand common phi/Data/CostPerTon_Ratios_ISSCAAP.csv',stringsAsFactors=F)

cost<-inner_join(costratio[,c('ISSCAAP.Group','Median','Mean')],costperton[,c('ISSCAAP.Group','Median','Mean')],by=c('ISSCAAP.Group'))

cost$Median.x<-format(round(cost$Median.x,2),nsmall=2)
cost$Mean.x<-format(round(cost$Mean.x,2),nsmall=2)

cost$Median.y<-format(round(cost$Median.y,0),nsmall=0)
cost$Mean.y<-format(round(cost$Mean.y,0),nsmall=0)

colnames(cost)<-c('ISSCAAP Category','Median Ratio','Mean Ratio','Median $/Ton','Mean $/Ton')

## Global results for fig 3
fig3<-read.csv('Results/6.0 global demand common phi/Data/Fig3data_Profit2050.csv',stringsAsFactors=F)

fig3$Policy<-c('Business As Usual S1','Business As Usual S2','RBFM S1','RBFM S2','FMSY S1','FMSY S2','Today')

fig3<-fig3[,c('Policy','Bio','Profit','Catch')]

fig3$Bio<-format(round(fig3$Bio,1),nsmall=2)
fig3$Profit<-format(round(fig3$Profit,1),nsmall=2)
fig3$Catch<-format(round(fig3$Catch,1),nsmall=2)

colnames(fig3)<-c('Policy','Biomass (MMT)','Annual Profit ($ Billions)','Harvest (MMT)')

## RAM stock list
ram<-read.csv('Results/6.0 global demand common phi/Data/Ram Stock List.csv',stringsAsFactors=F)

ram<-ram[,c('CommName','Country','RegionFAO')] %>%
  arrange(Country)

colnames(ram)<-c('Stock Name','Assigned Country','FAO Region')

## Commodity categories

commods<-read.csv('Data/CommodityNames.csv',stringsAsFactors=F)

  for(a in 1:nrow(commods))
  {
  if(commods$BestGuess...Not.included.in.2020[a]=='X') commods$CommodityName[a]<-paste(commods$CommodityName[a],'*',sep='')
  }

SpeciesCategoriesToOmit<- c('Corals','Frogs and other amphibians','Eared seals, hair seals, walruses',
                            'Turtles','Pearls, mother-of-pearl, shells','Crocodiles and alligators','Miscellaneous aquatic plants','Freshwater crustaceans','Sperm-whales, pilot-whales','Green seaweeds','Red seaweeds','Brown seaweeds','Sea-squirts and other tunicates','Blue-whales, fin-whales','Miscellaneous aquatic mammals','Sponges','Krill, planktonic crustaceans','Miscellaneous freshwater fishes','River eels','Freshwater molluscs')

commods<-commods %>%
  select(ISSCAAP,CommodityName) %>%
  filter(!(ISSCAAP %in% SpeciesCategoriesToOmit) & ISSCAAP!='Marine fishes nei') %>%
  arrange(CommodityName)

colnames(commods)<-c('Species Category (ISSCAAP)','Commodity Category')


```


**Table S5**  
Fishing costs as a fraction of revenue and per-ton catch. Results suggest that profitability varies considerably across ISSCAAP categories
```{r,echo=FALSE,results='asis'}
kable(cost)
```

**Table S6**  
Global results for all policies (BAU, RBFM, and FMSY) and scenarios (S1 and S2) in 2050 as presented in Figure 3 of the main tex 
```{r,echo=F,results='asis'}
kable(fig3)
```

**Table S7**  
Percent of upside in profit from NEI fisheries by Country for the RBFM policy applied to stocks of conservation concern
```{r,echo=FALSE,results='asis'}
kable(nei[,c('Country','Scenario','Policy','Percent of Country Profits from NEIs')])
```

**Table S8**  
Fisheries included from the RAM Legacy Stock Assessment Database. Upsides for fisheries listed as "Multinational" were distributed to countries identified as participating in that fishery using FAO landings data (SOM)
```{r,echo=F,results='asis'}
kable(ram)
```

**Table S9**  
ISSCAAP categories and their assigned commodity market categories as per Delgado et al. Asterisks indicate ISSCAAP categories that have been modified by the FAO since Delgado et al. and were categorized by the authors. ISSCAAP categories not included in Delgado et al. were categorized as 'Other marine animals'
```{r,echo=F,results='asis'}
kable(commods)
```


